MLH1 Data Report

### 2019-09-05 doc/Main_Report.rmd
updates to main report

Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents. Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.

@@

- Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.

-  Report basic statistics of results and to help keep track of the number of images for each category and mouse

*Main questions*

1. Which strains do I have enough mouse specific MLH1 data for?
2. What data will go into the FINAL dataset

@@

all cell image file names are anonimized and moved into bathces of a few hundred images.

Definition of passed quantified ect.

@@

*What were guidlines for quantification of lower quality cells / faint foci?*

-Count foci that look like other foci in the image.

@@

This section shows / displays / catches mice which haven't been stained or quantified yet.

@@

# Tests of dataset

Main metric is mean CO count per mouse. Assess the quality of this data with the following tests

@@

These aspects will be analyzed with a linear models on subsets of data (within category)

A general ancedote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (useing some of Beth's data)

There is a subset of data which has been counted mulitple times -- use that to test the repeatbility.

## Age Effect

This section is meant to double check the age distributions of male mice.

```{r age.hists, echo=FALSE, eval=FALSE}

meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

@@

 Is there an age effect?

make a subetted dataset for ploting mean mouse (quality < 5) MLH1 count and age.  Mice with enough ages:
WSB, KAZ, LEW, G, Cast, SPIC, MSM, MOLF, CZECH

```{r age.effect.model}
#run a age effect linear regression!

@@

summary(lm.me)

```{r age.effect.plot}

@@

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(15,39))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age Effect on CO mean")+facet_wrap(~subsp.y)
age.effect.plot
#todo add line to strain.y

@@

Given this plot it seems that there isn't a age effect. I should double check some of these mice

#### Materinal Age effect

@@

### litter effect?

## Spread and Stain Effects

*month dissected*

*staining batches*

## Room effect (envioronment)

@@

## cell number (power analysis results)

@@

@@

*Are the MLH1 distributions close to normal?*

```{r, MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
#consider making distribution with quality considered...

gg <- data.frame(weight=rep(x,nrow(df)),df)
gg$y <- with(gg,dnorm(x,mean,sd))
gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * diff(range(weight$weight))/30

@@

  geom_smooth(method = lm, se = FALSE)

  geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

@@

#if i'm testing normalge

#function = glm(y ~ x, family = poisson)
#add Poission curve behind
#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)

#not sure how to add Expected poisson curve

@@

How do these compare to Poission? (mean == variance)
I should make 'expected' distributions behind these.

@@

```{r, passing.mice, echo=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table
passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

@@

*Answer to questions*

@@

```{r, scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
#make lattice / plots seperated by strain and sex
#facet

@@

I ploted the mean of each quality bin with a red dot. From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data. This is most pronounced in MSM males and least pronounced in G males. The CAST female data is dummy data.

@@

*Is there mouse level MLH1 quality score bias?*

latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.

below are code chuncks that make a big dfs of CO.count by quality score plot. The code has been commented out, and moved to it's own doc.

@@

The mouse specific scatter plots aren't show because there are too many. (but they should be used for investigating specific mice).
Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse. The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.

PWD females and a very large range. This is driven by data from 17apr15_PWD_f4. These data should be looked at more closely.

In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality. This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.

@@

*-which strain has highest HetC?*

Start thinking of the direction for analysis

## within mouse

### female and male comparison

## within category (strain)

```{r HetC.plot, echo=FALSE,  fig.width=10, fig.height=5}
no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")

@@

### female and male comparison

## Subspecies level comparison

```{r mixed.model, echo=FALSE}

#build mixed model

#run simulations for accounting for sampling error

### Subspecies examples / Data

list the strains/data used 
Mouse average MLH1 counts,

### Male Specific results

Not all strains have female data. Here are results from all strains wih male data.

### female and male comparison

>>>>

Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice and related rodents. Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.

@@

- Display plots and figures of distributions of MLH1 counts to assess initial overall patterns.

-  Report basic statistics of results and to help keep track of the number of images for each category and mouse.

@@

all cell image file names are anonymized and moved into batches of a few hundred images.

@@

@@

This section shows the number of mice which haven't been stained or quantified yet.

@@

A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data). There is a subset of data which has been counted multiple times -- use that to test the repeatbility.

## Effects on the dataset

Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.

@@

These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.

### Age Effect

Some papers have documented age effects for male genome-wide recombination rate (cite).

The initial analysis uses plots. This should be double checked with a general linear model as well.

```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

@@

Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.

```{r age.effect.model, include=FALSE}
#run a age effect linear regression!

@@

#summary(lm.me)
```

```{r age.effect.plot, warning=FALSE, echo=FALSE}

@@

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(19,34))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age and mean CO per mouse")+facet_wrap(~subsp.y)
age.effect.plot
#todo add line to strain.y

@@

While many slopes are different than 0, I am not convinced there is a clear age effect. Most strains don't have more than 2 ages. I will double check some of these mice and check a general linear model.

### Maternal Age effect

Check if there is a maternal age effect for females and males.

@@

### Litter effect

Compare female spreads made on the same day, but from different mothers/litters.

### Spread and Stain Effects

> *month dissected*

For mice strains that have a similar MLH1 level (control for strain MLH1 effect), check if a month effect for MLH1 count (as seen in Peromyscus).

> *staining batches*

Compare observations from the same mouse, but different MLH1 stains.

### Room effect (envioronment)

@@

### Cell number (power analysis results)

Insert the code and results for effects of cell number/observations on MLH1 means.

@@

```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))

#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

#outgroups
#SPIC
HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
#unique(HetC.SPIC$mouse)
#non-passing mice
SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
#calq sex mean
SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]

#SPRET
HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
#unique(HetC.SPRET$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]

#CAROLI
HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
#unique(HetC.CAROLI$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

#cast
HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
#unique(HetC.CAST$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]

#HMI
HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
#unique(HetC.HMI$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

#dom

#WSB
HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
#unique(HetC.WSB$mouse)
#non-passing mice
WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
#calq sex mean
WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]

#LEW
HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
unique(HetC.LEW$mouse)
#non-passing mice
LEW.remove.mice <- c("4jan17_LEW_f2")
HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]

#calq sex mean
LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]

#G
HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
#unique(HetC.G$mouse)
#non-passing mice
G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]

#calq sex mean
G.mean.MLH1.obs <- ddply(.data=HetC.G, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]

#PERC
HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
#unique(HetC.PERC$mouse)
#non-passing mice
#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]

#calq sex mean
PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]

#MUSC

#PWD
HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
#unique(HetC.PWD$mouse)
#non-passing mice
PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]

# CZECH
HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
#unique(HetC.CZECH$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]

# SKIVE
HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
#unique(HetC.SKIVE$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]

# KAZ
HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
#unique(HetC.KAZ$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]

#TOM
HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
#unique(HetC.TOM$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]

#AST
HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
#unique(HetC.AST$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]

#Molosinus

#MSM
HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
unique(HetC.MSM$mouse)

#non-passing mice
MSM.remove.mice <- c("12sep16_MSM_f3")
HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]

#calq sex mean
MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]

#MOLF
HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
unique(HetC.MOLF$mouse)

#non-passing mice
MOLF.remove.mice <- c("27oct18_MOLF_m3")
HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]

#calq sex mean
MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]

> *-Which strain has highest HetC?*

```{r HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot

#make scatter plot for HetC
#color by strain

#set up data
#make female male plot

HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#
#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]

HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]

HetC_strain_table.M$sex <- "male"
HetC_strain_table.F$sex <- "female"

HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")

HetC_strain_table <- add_subsp(HetC_strain_table)

#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")

HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
#str(HetC_strain_table)

Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey"))

#musc - coral, Dom - cadetblue (grey blue),

Het.C.plot
#this plot is missing several

3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
1 of the musc strains has female biased heterochiasmy, KAZ.

@@

> *Are the MLH1 distributions close to normal?*

```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}

#gg <- data.frame(weight=rep(x,nrow(df)),df)
#gg$y <- with(gg,dnorm(x,mean,sd))
#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30

@@

#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

@@

#function = glm(y ~ x, family = poisson)
#add Poission curve behind
#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
#not sure how to add Expected poisson curve

@@

How do these compare to Poisson? (mean == variance)
I should make 'expected' distributions behind these.

@@

```{r passing.mice, echo=FALSE, warning=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table
passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

@@

@@

```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}

#make lattice / plots seperated by strain and sex
#facet

@@

I plotted the mean of each quality bin with a red dot. From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data. This is most pronounced in MSM males and least pronounced in G males. The CAST female data is dummy data.

@@

> *Is there mouse level MLH1 quality score bias?*

Below are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.

@@

The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.

Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse. The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.

@@

## Models for manuscript

Rough ideas for models of evolution of Heterochiasmy and current idea for how to test them.

- HetC arises due to more precursors (DSBs)

    - Compare DSB counts

- Larger cell volume drives longer chrms axis (and maybe more DSB)

    - Measure meiocyte diameter in males and females. Compare spermatocyte diameters in male biased and female biased strains. 
    - Compare DSBs counts and average SC/axis lengths.

- Difference in astral/MT organization

    - ?

- Something about interference and sister cohesin area

    - Develope metric for sister cohesin area, and look at initial patterns.

## Mixed Model, compare within and between subspecies patterns

```{r mixed.model, echo=FALSE}
#build mixed model
#run simulations for accounting for sampling error
# old code uses base plot

### Another Heterochiasmy Plot

```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average"}

no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")

@@
